### 2018-07-11 submission/manuscript.Rmd
add missing noun in first paragraph
fix figure legends

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a mock community [Figure 2E]. This total number of OTUs also varied by polymerase used for these mock community samples [Fiugre 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for both mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occured independently of the polymerase used [Figure 4]. Additionally, there were no differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing, the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between individuals.

@@

**Figure 1: The number and range of OTUs in fecal samples varies by polymerase.** The points represent the median number of OTUs of all four fecal samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The range in the number of OTUs detected in the different fecal samples increased as cycle number increased. This increased range also was larger for specific HiFi DNA polymerases.

**Figure 2: The number and range of OTUs in mock samples varies by polymerase.** The points represent the median number of OTUs for the mock samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The dotted black line represents the number of OTUs detected when only the references sequences for the mock community are clustered. The range in the number of OTUs detected in the mock samples increased as cycle number increased. This range was also larger for specific HiFi DNA polymerases.

**Figure 3: Lower five-cycle intervals have larger differences than higher five-cycle intervals using Bray-Curtis.** A) within person differences based on the next 5-cycle PCR interval in fecal samples. B) Within replicate difference based on the next 5-cycle PCR interval in mock samples. The points represent the median Bray-Curtis index for the samples. The lines represent the range of the minimum and maximum Bray-Curtis index value for each PCR 5-cycle increment comparison. The closer a sample is to a Bray-Curtis index of 1.00 the more dissimilar the bacterial community is of the two compared number of cycles.

**Figure 4: Variation in error rate by polymerase in mock community samples are similar across the number of cycles used.** The error bars represent the 75% interquartile range of the median error rate.

**Figure 5: Chimera prevalence varies by polymerase and increases with higher number of cycles used.** A) Percentage of chimeric sequences without the removal of chimeras with VSEARCH. C) Percentage of chimeric sequences with the removal of chimeras with VSEARCH. C) The total percent of chimeric sequences removed with VSEARCH by cycle number. The error bars represent the 75% interquartile range of the median.

**Figure 6: Regardless of polymerase used the higher the chimera prevalence the higher the observed number of OTUs in mock community samples.**

>>>>

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a mock community [Figure 2E]. This total number of OTUs also varied by polymerase used for these mock community samples [Fiugre 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for both mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occured independently of the polymerase used [Figure 4]. Additionally, there were no OTU-based differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing, the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between individuals.

@@

**Figure 1: Relative abundance differences due to GC is consistent across polymerases.** At each number of cycle the points and lines represent the median and with the minimum and maximum relative abundance respectively. Regardless of the number of cycle used a consistent difference between high and low GC bacteria was observed. The dotted line represents the actual relative abundance that each bacterium should be at within the mock community sample.

**Figure 2: Error rate and chimera prevalence vary by polymerase and affect the number of observed OTUs.** A) The error bars represent the 75% interquartile range of the median error rate. B) Percentage of chimeric sequences without the removal of chimeras with VSEARCH. C) Percentage of chimeric sequences with the removal of chimeras with VSEARCH. D) The total percent of chimeric sequences removed with VSEARCH by cycle number. The error bars represent the 75% interquartile range of the median. E) Chimera prevalence and the the observed number of OTUs are strongly correlated in mock community samples.

**Figure 3: The number and range of OTUs in mock and fecal samples varies by polymerase.** The points represent the median number of OTUs for the mock samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The dotted black line represents the number of OTUs detected when only the references sequences for the mock community are clustered. A) The range in the number of OTUs detected in the mock samples increased as cycle number increased. This range was also larger for specific HiFi DNA polymerases. B) The range in the number of OTUs detected in the different fecal samples increased as cycle number increased. This increased range also was larger for specific HiFi DNA polymerases.

**Figure 4: Lower five-cycle intervals have larger differences than higher five-cycle intervals using Bray-Curtis.** A) Within replicate difference based on the next 5-cycle PCR interval in mock samples. B) Within person differences based on the next 5-cycle PCR interval in fecal samples. The points represent the median Bray-Curtis index for the samples. The lines represent the range of the minimum and maximum Bray-Curtis index value for each PCR 5-cycle increment comparison. The closer a sample is to a Bray-Curtis index of 1.00 the more dissimilar the bacterial community is of the two compared number of cycles.

**Figure 5: Biological variation in bacterial community composition due to individual was larger than number of cycles and polymerase used.** A) Using 16S rRNA gene sequencing, the classification of individuals was higher than any other model tested. The lines represent the median probability of correct classification and the gray points represent the individual probability from one run out of the 100 10-fold CV runs. B) The most important OTUs to the model classifying individuals are not the same as those that are most important for the number of cycles or polymerase used.

### 2018-07-11 submission/manuscript.tex
add missing noun in first paragraph
fix figure legends

Previous studies have reported GC dependent amplification differences
(Polz \& Cavanaugh, 1998) and we confirm that these differences still
persist in high fidelity DNA polymerase when analyzing mock community
samples {[}Figure 1{]}. Additionally, these specific polymerases have
different sequence error rates and chimera prevalence {[}Figure 2{]}.
Chimera prevalence in particular was strongly associated with the total
number of OTUs detected in a mock community {[}Figure 2E{]}. This total
number of OTUs also varied by polymerase used for these mock community
samples {[}Fiugre 3A{]}. The range in the number of OTUs observed
increased with the number of cycles and the change in range was
dependent on the polymerase for both mock community and fecal samples
{[}Figure 3{]}. However, these differences did not translate to large
bacterial community composition changes based on polymerase {[}Figure
4{]}. Even though there were bacterial community composition changes
based on the number of cycles they occured independently of the
polymerase used {[}Figure 4{]}. Additionally, there were no differences
observed within the number of cycles between any of the polymerases for
fecal samples. When Random Forest models were created to classify number
of cycles, polymerase, and individual based on the 16S rRNA gene
sequencing, the model performance for individuals was significantly
greater than all other models tested {[}Figure 5A{]}. In addition to the
better model performance of individuals, the most important OTUs to this
model were not the most important OTUs for the other models tested
{[}Figure 5B{]}. Our observations show that differences between high
fidelity DNA polymerases can be found. However, when using 16S rRNA gene
sequencing these technical differences are dwarfed by biological
differences between individuals.

@@

\textbf{Figure 1: The number and range of OTUs in fecal samples varies
by polymerase.} The points represent the median number of OTUs of all
four fecal samples. The lines represent the range of the minimum and
maximum number of OTUs detected within the four fecal samples. The range
in the number of OTUs detected in the different fecal samples increased
as cycle number increased. This increased range also was larger for
specific HiFi DNA polymerases.

\textbf{Figure 2: The number and range of OTUs in mock samples varies by
polymerase.} The points represent the median number of OTUs for the mock
samples. The lines represent the range of the minimum and maximum number
of OTUs detected within the four fecal samples. The dotted black line
represents the number of OTUs detected when only the references
sequences for the mock community are clustered. The range in the number
of OTUs detected in the mock samples increased as cycle number
increased. This range was also larger for specific HiFi DNA polymerases.

\textbf{Figure 3: Lower five-cycle intervals have larger differences
than higher five-cycle intervals using Bray-Curtis.} A) within person
differences based on the next 5-cycle PCR interval in fecal samples. B)
Within replicate difference based on the next 5-cycle PCR interval in
mock samples. The points represent the median Bray-Curtis index for the
samples. The lines represent the range of the minimum and maximum
Bray-Curtis index value for each PCR 5-cycle increment comparison. The
closer a sample is to a Bray-Curtis index of 1.00 the more dissimilar
the bacterial community is of the two compared number of cycles.

\textbf{Figure 4: Variation in error rate by polymerase in mock
community samples are similar across the number of cycles used.} The
error bars represent the 75\% interquartile range of the median error
rate.

\textbf{Figure 5: Chimera prevalence varies by polymerase and increases
with higher number of cycles used.} A) Percentage of chimeric sequences
without the removal of chimeras with VSEARCH. C) Percentage of chimeric
sequences with the removal of chimeras with VSEARCH. C) The total
percent of chimeric sequences removed with VSEARCH by cycle number. The
error bars represent the 75\% interquartile range of the median.

\textbf{Figure 6: Regardless of polymerase used the higher the chimera
prevalence the higher the observed number of OTUs in mock community
samples.}

>>>>

Previous studies have reported GC dependent amplification differences
(Polz \& Cavanaugh, 1998) and we confirm that these differences still
persist in high fidelity DNA polymerase when analyzing mock community
samples {[}Figure 1{]}. Additionally, these specific polymerases have
different sequence error rates and chimera prevalence {[}Figure 2{]}.
Chimera prevalence in particular was strongly associated with the total
number of OTUs detected in a mock community {[}Figure 2E{]}. This total
number of OTUs also varied by polymerase used for these mock community
samples {[}Fiugre 3A{]}. The range in the number of OTUs observed
increased with the number of cycles and the change in range was
dependent on the polymerase for both mock community and fecal samples
{[}Figure 3{]}. However, these differences did not translate to large
bacterial community composition changes based on polymerase {[}Figure
4{]}. Even though there were bacterial community composition changes
based on the number of cycles they occured independently of the
polymerase used {[}Figure 4{]}. Additionally, there were no OTU-based
differences observed within the number of cycles between any of the
polymerases for fecal samples. When Random Forest models were created to
classify number of cycles, polymerase, and individual based on the 16S
rRNA gene sequencing, the model performance for individuals was
significantly greater than all other models tested {[}Figure 5A{]}. In
addition to the better model performance of individuals, the most
important OTUs to this model were not the most important OTUs for the
other models tested {[}Figure 5B{]}. Our observations show that
differences between high fidelity DNA polymerases can be found. However,
when using 16S rRNA gene sequencing these technical differences are
dwarfed by biological differences between individuals.

@@

\textbf{Figure 1: Relative abundance differences due to GC is consistent
across polymerases.} At each number of cycle the points and lines
represent the median and with the minimum and maximum relative abundance
respectively. Regardless of the number of cycle used a consistent
difference between high and low GC bacteria was observed. The dotted
line represents the actual relative abundance that each bacterium should
be at within the mock community sample.

\textbf{Figure 2: Error rate and chimera prevalence vary by polymerase
and affect the number of observed OTUs.} A) The error bars represent the
75\% interquartile range of the median error rate. B) Percentage of
chimeric sequences without the removal of chimeras with VSEARCH. C)
Percentage of chimeric sequences with the removal of chimeras with
VSEARCH. D) The total percent of chimeric sequences removed with VSEARCH
by cycle number. The error bars represent the 75\% interquartile range
of the median. E) Chimera prevalence and the the observed number of OTUs
are strongly correlated in mock community samples.

\textbf{Figure 3: The number and range of OTUs in mock and fecal samples
varies by polymerase.} The points represent the median number of OTUs
for the mock samples. The lines represent the range of the minimum and
maximum number of OTUs detected within the four fecal samples. The
dotted black line represents the number of OTUs detected when only the
references sequences for the mock community are clustered. A) The range
in the number of OTUs detected in the mock samples increased as cycle
number increased. This range was also larger for specific HiFi DNA
polymerases. B) The range in the number of OTUs detected in the
different fecal samples increased as cycle number increased. This
increased range also was larger for specific HiFi DNA polymerases.

\textbf{Figure 4: Lower five-cycle intervals have larger differences
than higher five-cycle intervals using Bray-Curtis.} A) Within replicate
difference based on the next 5-cycle PCR interval in mock samples. B)
Within person differences based on the next 5-cycle PCR interval in
fecal samples. The points represent the median Bray-Curtis index for the
samples. The lines represent the range of the minimum and maximum
Bray-Curtis index value for each PCR 5-cycle increment comparison. The
closer a sample is to a Bray-Curtis index of 1.00 the more dissimilar
the bacterial community is of the two compared number of cycles.

\textbf{Figure 5: Biological variation in bacterial community
composition due to individual was larger than number of cycles and
polymerase used.} A) Using 16S rRNA gene sequencing, the classification
of individuals was higher than any other model tested. The lines
represent the median probability of correct classification and the gray
points represent the individual probability from one run out of the 100
10-fold CV runs. B) The most important OTUs to the model classifying
individuals are not the same as those that are most important for the
number of cycles or polymerase used.
